---
title: " "
author: " "
email: " "
date: " "
output: 
  bookdown::html_document2:
    code_folding: hide
    toc: TRUE
    toc_depth: 4
    toc_float: TRUE
    collapsed: FALSE
    theme: spacelab
    fig_caption: TRUE
    number_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = F,
	message = FALSE,
	warning = FALSE,
	dpi = 300,
	progress = FALSE,
	cache = FALSE,
	fig.align="center",
	fig.width=7
)


set.seed(513)
options(repos = c(CRAN = "http://cran.rstudio.com"))
options(timeout=60*30) #timeout downloads that last longer than 30 minutes

if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  here, # easy file structure
  tidyverse, # data manipulation
  tidymodels, 
  reshape, # old data manipulation pkg
  ggplot2, # figures
  patchwork, # multiple figures
  grid, # multiple figures
  ggsci, # color palettes
  colorspace, # color palettes
  RColorBrewer, # color palettes
  knitr, # markdown documents
  kableExtra, # tables in markdown
  flextable, # other table options
  bookdown, # figure numbering in markdown
  sf, # spatial data (new pkg)
  terra, # raster data
  raster, # (old pkg)
  tmap, # easy plotting of maps
  terrainr, # elevation data
  spatialEco, # hli, tpi
  sdm,# species distribution modeling
  rgdal,
  parallel,
  FNN,
  maptools,
  rgeos,
  elevatr
) 

cores <- parallel::detectCores() # Set number of cores for parallel processing

# Set custom plotting theme
theme_new <- function(base_size = 9,base_family = "Helvetica"){
  theme_classic(base_size = base_size, base_family = base_family) %+replace%
    theme(
      axis.line.x = element_line(color="black", size = 0.25),
      axis.line.y = element_line(color="black", size = 0.25),
      axis.title = element_text(size = 9),
      axis.text = element_text(colour="black", size=8),
      legend.key=element_rect(colour=NA, fill =NA),
      panel.grid = element_blank(),   
      plot.background = element_rect(fill = NA, colour = NA),
      panel.border = element_rect(fill = NA, colour = NA),
      panel.background = element_rect(fill = "white", colour = "black"), 
      strip.background = element_rect(fill = "white"),
      strip.text = element_text(size = 9)
      
    )
}
theme_set(theme_new())


# Set directory structure for project
dir.create(here("Data"), showWarnings = FALSE)
dir.create(here("Data", "Spatial"), showWarnings = FALSE)
dir.create(here("Results"), showWarnings = FALSE)
dir.create(here("Results", "Figures"), showWarnings = FALSE)
```

# Objectives:

JFSP Aim 4: Produce maps of current and future aspen habitat suitability

### Proposed methods

To quantify aspen habitat suitability, we will construct statistical models of aspen presence that incorporate historical climate, soil, and topography variables as predictors. We will randomly sample the Sentinel-2 based maps of aspen produced here (Aim 1) to generate 1000 aspen presences and absences within the SRM. To limit the influence of spatial autocorrelation (Veloz 2009), we will filter all the records by a minimum distance of 1 km. For each presence and absence record, we will then characterize the historical climate, topography and soils conditions. We will characterize fine scale variability in incident radiation using the heat load index, which we will calculate from 10 m DEM. To characterize soil characteristics that may influence habitat suitability for aspen, we will use 30 m grids of soil pH and percent clay cover (Chaney et al. 2019). To characterize historical (1981-2015) climate, we will use climate data from gridMet (Abatzoglou 2013). Specifically, we will obtain mean total winter, summer, and annual precipitation and average summer and winter temperatures, which are related to aspen presence (Rehfeldt et al. 2009, Ellenwood et al. 2015). Additionally, we will calculate mean annual actual evapotranspiration (AET) and climatic water deficit (CWD), two integrative measures of the amount of energy and water available for plant growth that are linked to plant species distributions (Stephenson 1990, Lutz et al. 2010). gridMET products are available at a ca. 4 km resolution, a scale that is too coarse to describe topographically controlled climate variation characteristic of mountainous environments (Franklin et al. 2013). To this end, we will downscale climate data to a 250 m resolution using gradient and inverse distance squared (GIDS) interpolation (Nalder and Wein 1998, Flint and Flint 2012), following methods outlined in Rodman et al. (2020)

We will then use an ensemble modeling approach to generate predictions of aspen habit suitability. We will model presence/absence using several machine learning approaches widely used in habitat suitability modeling (e.g. Bayesian Additive Regression Trees Chipman et al. 2010). Individual models will first be tuned using the R package caret (Kuhn 2020) and then we will use cross-validation to estimate the performance of all models and generate ensemble predictions using the R package SuperLearner (Polley et al. 2019). To generate future predictions of aspen habitat suitability, we will incorporate future (2021-2099) climate data generated using the Multivariate Adaptive Constructed Analogs (MACA) method (Abatzoglou and Brown 2012). We will obtain future climate projections under an intermediate pathway (RCP4.5) and a high emissions pathway (RCP8.5) (IPCC 2014). To encompass the variability in global circulation model (GCM) projections, we will select three GCMs that span the range of project change. We will then downscale future climate data to 250 m following the methods described above. Finally, we will use fitted models to make projections of the future of aspen habitat suitability.


# Data
## Download data
### Ecoregions
```{r download ecoregions}
temp <- here("Data", "Spatial", "Ecoregions","us_eco_l3.zip")
if(!file.exists(temp)){
download.file("https://gaftp.epa.gov/EPADataCommons/ORD/Ecoregions/us/us_eco_l3.zip", temp, mode="wb")
unzip(temp, exdir=here("Data", "Spatial", "Ecoregions"))
}
```

### High resolution aspen cover
Cook and Balch (unpublished)
- This code will download a 10 m resolution map of aspen cover for the Southern Rocky Mountain Ecoregion, which extends from southern Wyoming to northern New Mexico. The map represents aspen cover in ca. 2019 and was produced from Sentinel imagery (JFSP AIM 1)
Updated aspen data with 0.64 probability

```{r}
dir.create(here("Data", "Spatial", "Aspen"), showWarnings = FALSE)
temp <- here("Data", "Spatial", "Aspen", "SRM_ASPEN_CLASSIFIED_RF_10M_V1_0223.tif", mode="wb")
if(!file.exists(temp)){
download.file("https://1drv.ms/u/s!Aq4s9rj0qHBilL4B4PH1amvdUoiLyg?e=jHRUWI", temp )
}
```


### Soils
```{r}
if(!file.exists(here("Data","Spatial", "Soils"))){
  dir.create(here("Data","Spatial", "Soils"), showWarnings = FALSE)

  #SRM %>% st_transform(crs=4326) %>% st_bbox --> lat: 35-43; lon:103-110
  lons <- c("110-109", "109-108", "108-107", "107-106", "106-105", "105-104", "104-103")
  lats <- c("3536", "3637", "3738", "3839","3940", "4041", "4142", "4243")

  # download mean pH of top 5-15 cm
  for(j in lons){
    for(k in lats){
      temp <- paste0(here("Data","Spatial", "Soils", "lat"), k,"_lon-", j, "_meanPh_5_15.tif")
      download.file(paste0("http://hydrology.cee.duke.edu/POLARIS/PROPERTIES/v1.0/ph/mean/5_15/lat",   k, "_lon-", j, ".tif"), temp, mode="wb")
    }
  }
  
  # Merge mean pH files
  ph.files <- list.files(path=here("Data","Spatial", "Soils"), pattern="meanPh_5_15.tif", full.names=T)
  ph.rast <- lapply(ph.files,FUN=rast)
  ph.rast <- sprc(ph.rast)
  ph.merge <- merge(ph.rast, first=T)
  writeRaster(ph.merge, here("Data","Spatial", "Soils", "meanPh_5_15-SRM.tif"))
  
  # download mean percent clay of top 5-15 cm
  for(j in lons){
    for(k in lats){
      temp <- paste0(here("Data","Spatial", "Soils", "lat"), k,"_lon-", j, "_meanClay_5_15.tif")
      download.file(paste0("http://hydrology.cee.duke.edu/POLARIS/PROPERTIES/v1.0/clay/mean/5_15/lat"  , k, "_lon-", j, ".tif"), temp, mode="wb")
    }
  }
  
  # Merge mean clay files
  clay.files <- list.files(path=here("Data", "Spatial", "Soils"), pattern="meanClay_5_15.tif",   full.names=T)
  clay.rast <- lapply(clay.files,FUN=rast)
  clay.rast <- sprc(clay.rast)
  clay.merge <- merge(clay.rast, first=T)
  writeRaster(clay.merge, here("Data", "Spatial", "Soils", "meanClay_5_15-SRM.tif"))
}
### download mean percent theta_s (saturated soil water content) of top 5-15 cm
  for(j in lons){
    for(k in lats){
      temp <- paste0(here("Data","Spatial", "Soils", "lat"), k,"_lon-", j, "_meantheta_s_5_15.tif")
      download.file(paste0("http://hydrology.cee.duke.edu/POLARIS/PROPERTIES/v1.0/theta_s/mean/5_15/lat"  , k, "_lon-", j, ".tif"), temp, mode="wb")
    }
  }
  
  # Merge theta_s files
  theta_s.files <- list.files(path=here("Data", "Spatial", "Soils"), pattern="meantheta_s_5_15.tif",   full.names=T)
  theta_s.rast <- lapply(theta_s.files,FUN=rast)
  theta_s.rast <- sprc(theta_s.rast)
  theta_s.merge <- merge(theta_s.rast, first=T)
  writeRaster(theta_s.merge, here("Data", "Spatial", "Soils", "meantheta_s_5_15-SRM.tif"))
  
  ##download mean percent om (organic matter) of top 5-15 cm
  for(j in lons){
    for(k in lats){
      temp <- paste0(here("Data","Spatial", "Soils", "lat"), k,"_lon-", j, "_meanom_5_15.tif")
      download.file(paste0("http://hydrology.cee.duke.edu/POLARIS/PROPERTIES/v1.0/om/mean/5_15/lat"  , k, "_lon-", j, ".tif"), temp, mode="wb")
    }
  }
  
 # Merge om files
  om.files <- list.files(path=here("Data", "Spatial", "Soils"), pattern="meanom_5_15.tif",   full.names=T)
  om.rast <- lapply(om.files,FUN=rast)
  om.rast <- sprc(om.rast)
  om.merge <- merge(om.rast, first=T)
  writeRaster(om.merge, here("Data", "Spatial", "Soils", "meanom_5_15-SRM.tif"))  
```

### DEM
```{r}
if(!file.exists(here("Data", "Spatial","DEM"))){
  
  dir.create(here("Data", "Spatial","DEM"), showWarnings = FALSE)
  
  SRM <- st_read(here("Data", "Spatial", "Ecoregions",
"us_eco_l3.shp")) %>% filter(US_L3NAME == 'Southern Rockies') %>% st_transform(crs=5071) %>% st_buffer(20000)

  # 30 m 
  dem <- get_tiles(data=SRM, output_prefix = here("Data", "Spatial", "DEM", "DEM30"), services="elevation",   resolution=30)
dem.rast <- lapply(dem$elevation,FUN=rast)
  for(j in 2:length(dem.rast)){dem.rast[[j]] <-  terra::project(dem.rast[[j]], dem.rast[[1]], align=T)}
  dem.sprc <- sprc(dem.rast)
  dem.mosaic <- mosaic(dem.sprc)
  writeRaster(dem.mosaic, here("Data", "Spatial", "DEM", "DEM30-mosaic.tif"), overwrite=T)
  
  # 250 m 
  dem <- get_tiles(data=SRM, output_prefix = here("Data", "Spatial", "DEM", "DEM250"), services="elevation",   resolution=250)
  dem.rast <- lapply(dem$elevation,FUN=rast)
  writeRaster(dem.rast[[1]], here("Data", "Spatial", "DEM", "DEM250.tif"), overwrite=T)
}
```

### Climate data
#### ClimateNA

```{r}
dir.create(here("Data","Spatial", "Climate"))
if(!file.exists(here("Data","Spatial", "Climate", "ClimateNA"))){
  dir.create(here("Data","Spatial", "Climate", "ClimateNA"), showWarnings = FALSE)
  temp <- here("Data","Spatial", "Climate", "ClimateNA", "Normal_1981_2010_bioclim.zip")
  download.file("https://s3-us-west-2.amazonaws.com/www.cacpd.org/CMIP6v73/normals/Normal_1981_2010_bioclim.zip", temp, mode="wb")
  unzip(temp, exdir=here("Data", "Spatial", "Climate", "CLimateNA"))
  
  ## 2011-2040 Emission Scenario SSP2-4.5
  temp <- here("Data","Spatial", "Climate", "ClimateNA","ensemble_8GCMs_ssp245_2011_2040_bioclim.zip")
  download.file("https://s3-us-west-2.amazonaws.com/www.cacpd.org/CMIP6v73/ensembles/ensemble_8GCMs_ssp245_2011_2040_bioclim.zip", temp, mode="wb")
  unzip(temp, exdir=here("Data", "Spatial", "Climate", "CLimateNA"))
  
  ## 2040-2070 Emission Scenario SSP2-4.5
  temp <- here("Data","Spatial", "Climate", "ClimateNA","ensemble_8GCMs_ssp245_2041_2070_bioclim.zip")
  download.file("https://s3-us-west-2.amazonaws.com/www.cacpd.org/CMIP6v73/ensembles/ensemble_8GCMs_ssp245_2041_2070_bioclim.zip", temp, mode="wb")
  unzip(temp, exdir=here("Data", "Spatial", "Climate", "CLimateNA"))
  
  ## 2071-2100 Emission Scenario SSP2-4.5
  temp <- here("Data","Spatial", "Climate", "ClimateNA","ensemble_8GCMs_ssp245_2071_2100_bioclim.zip")
  download.file("https://s3-us-west-2.amazonaws.com/www.cacpd.org/CMIP6v73/ensembles/ensemble_8GCMs_ssp245_2071_2100_bioclim.zip", temp, mode="wb")
  unzip(temp, exdir=here("Data", "Spatial", "Climate", "CLimateNA"))
 
  ##2011-2040 Emission Scenario SSP5-8.5
  temp <- here("Data","Spatial", "Climate", "ClimateNA","ensemble_8GCMs_ssp585_2011_2040_bioclim.zip")
  download.file("https://s3-us-west-2.amazonaws.com/www.cacpd.org/CMIP6v73/ensembles/ensemble_8GCMs_ssp585_2011_2040_bioclim.zip", temp, mode="wb")
  unzip(temp, exdir=here("Data", "Spatial", "Climate", "CLimateNA"))
 
  ##2041-2070 Emission Scenario SSP5-8.5
  temp <- here("Data","Spatial", "Climate", "ClimateNA","ensemble_8GCMs_ssp585_2041_2070_bioclim.zip")
  download.file("https://s3-us-west-2.amazonaws.com/www.cacpd.org/CMIP6v73/ensembles/ensemble_8GCMs_ssp585_2041_2070_bioclim.zip", temp, mode="wb")
  unzip(temp, exdir=here("Data", "Spatial", "Climate", "CLimateNA"))
  
  ##2071-2100 Emission Scenario SSP5-8.5
  temp <- here("Data","Spatial", "Climate", "ClimateNA","ensemble_8GCMs_ssp585_2071_2100_bioclim.zip")
  download.file("https://s3-us-west-2.amazonaws.com/www.cacpd.org/CMIP6v73/ensembles/ensemble_8GCMs_ssp585_2071_2100_bioclim.zip", temp, mode="wb")
  unzip(temp, exdir=here("Data", "Spatial", "Climate", "CLimateNA"))
}
```

#### TopoFire

```{r}
dir.create(here("Data","Spatial", "Climate"))
if(!file.exists(here("Data","Spatial", "Climate", "TopoFire"))){
  dir.create(here("Data","Spatial", "Climate", "TopoFire"), showWarnings = FALSE)
  temp <- here("Data","Spatial", "Climate", "TopoFire", "aet_1981-2010_mean.tif")
  download.file("https://topofire.dbs.umt.edu/public_data/topofire_weather/normals_old/aet_1981-2010_mean.tif", temp, mode="wb")
  temp <- here("Data","Spatial", "Climate", "TopoFire", "def_1981-2010_mean.tif")
  download.file("https://topofire.dbs.umt.edu/public_data/topofire_weather/normals_old/def_1981-2010_annual_mean.tif", temp, mode="wb")
}
```

## Data processing
### Calculate heat load index
```{r}
if(!file.exists(here("Data","Spatial", "DEM", "DEM30-mosaic-HLI.tif"))){
  dem.mosaic <- rast(here(here("Data","Spatial", "DEM", "DEM30-mosaic.tif")))
  hli.rast <- hli(dem.mosaic, force.hemisphere = "northern")
  writeRaster(hli.rast, here("Data","Spatial", "DEM", "DEM30-mosaic-HLI.tif"))
}
```

### Caculate topographic position index (TPI)
#### 3 cell neighborhood
```{r}
if(!file.exists(here("Data","Spatial", "DEM", "DEM30-mosaic-TPI3.tif"))){
  dem.mosaic <- rast(here(here("Data","Spatial", "DEM", "DEM30-mosaic.tif")))
  tpi.rast <- tpi(dem.mosaic, scale=3)
  writeRaster(tpi.rast, here("Data","Spatial", "DEM", "DEM30-mosaic-TPI3.tif"))
}
```

#### 15 cell neighborhood
```{r}
if(!file.exists(here("Data","Spatial", "DEM", "DEM30-mosaic-TPI15.tif"))){
  dem.mosaic <- rast(here(here("Data","Spatial", "DEM", "DEM30-mosaic.tif")))
  tpi.rast <- tpi(dem.mosaic, scale=15)
  writeRaster(tpi.rast, here("Data","Spatial", "DEM", "DEM30-mosaic-TPI15.tif"))
}
```

## Downscaling


```{r, downscale}
source(here("Code", "GIDS-Downscaling-SJH.R"))

# Set desired projection
proj.proj <- "+proj=utm +zone=13 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"

# Import study area boundary
study.area <- st_read(here("Data", "Spatial", "Ecoregions", "us_eco_l3.shp")) %>% filter(US_L3NAME == 'Southern Rockies') %>% st_transform(crs(proj.proj)) %>% st_buffer(20000)

# Import template for downscaling
template <- rast(here("Data", "Spatial", "DEM", "DEM250.tif"))
template <- template %>% terra::project(proj.proj)
template<- template %>% crop(study.area)
 
# Warning message happens because number of raster cells is not evenly divisible
# by number of cores on PC

select.climate.variables <-c("CMD", "DD_0", "DD5", "MAP", "MCMT", "MSP", "MWMT", "RH", "SHM", "TD","Tmax")
dir.create(here("Data", "Spatial", "Downscaled"), showWarnings = FALSE)
for(j in select.climate.variables){
  if(!file.exists(paste0(here("Data", "Spatial", "Downscaled", "ensemble_8GCMs_ssp245_2011_2040_"), j, "-downscaled.tif"))){
    ptm <- proc.time()
    clim <- rast(paste0(here("Data", "Spatial", "Climate", "CLimateNA", "ensemble_8GCMs_ssp245_2011_2040", "ensemble_8GCMs_ssp245_2011_2040_bioclim", "ensemble_8GCMs_ssp245_2011_2040_"), j, ".tif"))
    clim <- clim %>% terra::project(proj.proj)
    clim <- clim %>% crop(study.area)
  
    if(j ==select.climate.variables[1]){template_coarse <- terra::project(template ,clim, method="bilinear")}   ## Creating second DEM at same resolution as climate grid; only do this for the first iteration
    
    ds_layer <- multiLevelInterpParrallel(boundary = study.area, ds_layer = clim,
                                        ancillary_list = list(template_coarse, template), 
                                        clip_dists = c(20000, 5000, 500))
  
    writeRaster(ds_layer, paste0(here("Data", "Spatial", "Downscaled", "ensemble_8GCMs_ssp245_2011_2040_"), j, "-downscaled.tif"))
    rm(ds_layer) # remove 
    gc() # clean up garbage (sometimes this helps)
    print(j) # print j so we know how far along our code is
    time <- proc.time() - ptm
    print(time[3]/60) # print time for each iteration 
  }

}

for(j in select.climate.variables){
  if(!file.exists(paste0(here("Data", "Spatial", "Downscaled", "ensemble_8GCMs_ssp245_2041_2070_"), j, "-downscaled.tif"))){

    ptm <- proc.time()
    clim <- rast(paste0(here("Data", "Spatial", "Climate", "CLimateNA", "ensemble_8GCMs_ssp245_2041_2070", "ensemble_8GCMs_ssp245_2041_2070_bioclim", "ensemble_8GCMs_ssp245_2041_2070_"), j, ".tif"))
    clim <- clim %>% terra::project(proj.proj)
    clim <- clim %>% crop(study.area)
  
    if(j ==select.climate.variables[1]){template_coarse <- terra::project(template,clim, method="bilinear")}   ## Creating second DEM at same resolution as climate grid; only do this for the first iteration
    
    ds_layer <- multiLevelInterpParrallel(boundary = study.area, ds_layer = clim,
                                        ancillary_list = list(template_coarse, template), 
                                        clip_dists = c(20000, 5000, 500))
    writeRaster(ds_layer, paste0(here("Data", "Spatial", "Downscaled", "ensemble_8GCMs_ssp245_2041_2070_"), j, "-downscaled.tif"))
    rm(ds_layer) # remove 
    gc() # clean up garbage (sometimes this helps)
    print(j) # print j so we know how far along our code is
    time <- proc.time() - ptm
    print(time[3]/60) # print time for each iteration
  }
}

for(j in select.climate.variables){
  if(!file.exists(paste0(here("Data", "Spatial", "Downscaled", "ensemble_8GCMs_ssp245_2071_2100_"), j, "-downscaled.tif"))){

    ptm <- proc.time()
    clim <- rast(paste0(here("Data", "Spatial", "Climate", "CLimateNA", "ensemble_8GCMs_ssp245_2071_2100", "ensemble_8GCMs_ssp245_2071_2100_bioclim", "ensemble_8GCMs_ssp245_2071_2100_"), j, ".tif"))
    clim <- clim %>% terra::project("+proj=utm +zone=13 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")
    clim <- clim %>% crop(study.area)
    
    if(j ==select.climate.variables[1]){template_coarse <- terra::project(template,clim, method="bilinear")}   ## Creating second DEM at same resolution as climate grid; only do this for the first iteration
    
    ds_layer <- multiLevelInterpParrallel(boundary = study.area, ds_layer = clim,
                                        ancillary_list = list(template_coarse, template), 
                                        clip_dists = c(20000, 5000, 500))
    writeRaster(ds_layer, paste0(here("Data", "Spatial", "Downscaled", "ensemble_8GCMs_ssp245_2071_2100_"), j, "-downscaled.tif"))
    rm(ds_layer) # remove 
    gc() # clean up garbage (sometimes this helps)
    print(j) # print j so we know how far along our code is
    time <- proc.time() - ptm
    print(time[3]/60) # print time for each iteration 
  }
}
```


## Analyses
#### Import data
```{r}
# RESAMPLE 
aspen <- rast(here("Data", "Spatial", "Aspen", "SRM_ASPEN_CLASSIFIED_RF_10M_V1_0522.tif"))
tpi3 <- rast(here("Data", "Spatial", "DEM", "DEM30-mosaic-TPI3.tif")) 
tpi15 <- rast(here("Data", "Spatial", "DEM", "DEM30-mosaic-TPI15.tif"))
hli <- rast(here("Data", "Spatial", "DEM", "DEM30-mosaic-HLI.tif"))
topo <- c(tpi3, tpi15, hli) %>% project(crs(aspen))
names(topo) <- c("tpi3", "tpi15", "hli")
if(!file.exists(filename=here("Data", "Spatial", "DEM", "Topo10.tif"))){
  topo.re <- resample(topo, aspen, method="near", threads=T, filename=here("Data", "Spatial", "DEM", "Topo10.tif"),overwrite=TRUE)
}
ph <- rast(here("Data", "Spatial", "Soils", "meanPh_0_5-SRM.tif"))  %>% project(crs(aspen))
clay <- rast(here("Data", "Spatial", "Soils", "meanClay_0_5-SRM.tif"))  %>% project(crs(aspen))
soil <- c(ph, clay)
names(soil) <- c("pH", "clay")
soil.re <- resample(soil, aspen, method="near", threads=T, filename=here("Data", "Spatial", "Soils", "Topo10.tif"),overwrite=TRUE)

climate <- list.files(here("Data", "Spatial", "Downscaled"), pattern="Normal", full.names=T)
climate <- rast(climate) %>% project(crs(aspen))
names(climate) <- sapply(sapply(sapply(sapply(list.files(here("Data", "Spatial", "Downscaled"), pattern="Normal", full.names=T), strsplit, split="2010_"), "[", 2), strsplit, split="-downscaled.tif"), "[", 1)
climate.re <- resample(climate, aspen, method="near", threads=T, filename=here("Data", "Spatial", "Downscaled", "Climate10.tif"),overwrite=TRUE)

predictors <- c(climate.re, soil.re, topo.re)
```

#random sampling of Aspen presence and absence
#Projecting all to WGS_1984 is importnat becasue biomod doesn't work in NAD..
#Correlation matrix and predictor variable selection
#BioMod2
```{r}
library(biomod2)
library(raster)
library(rasterVis)
library(ggplot2)
library(gridExtra)
library(dplyr)#for correlation matrix
library(gridExtra)
library(usdm)#for VIF
######Sampling Aspen Data from Max Data
library(rgdal)#required for random numbers of raw selection
setwd("D:/BioMod2Try/Max_Aspen_Raster/samplingjune42023")
myshapefile <- readOGR(dsn = ".",layer = "presence1")

myshapefile
class(myshapefile)
my_df<-as.data.frame(myshapefile)
dim(my_df)
my_sample<-my_df[sample(nrow(my_df),2000),]
write.csv(my_sample,"D:/BioMod2Try/Max_Aspen_Raster/samplingjune42023/presence.csv")
#absence
setwd("D:/BioMod2Try/Max_Aspen_Raster/samplingjune42023")
myshapefile <- readOGR(dsn = ".",layer = "absence0")
myshapefile
class(myshapefile)
my_df<-as.data.frame(myshapefile)
dim(my_df)
my_sample<-my_df[sample(nrow(my_df),2000),]
write.csv(my_sample,"D:/BioMod2Try/Max_Aspen_Raster/samplingjune42023/absence.csv")

#Stacking predictor variables data
setwd("D:/BioMod2Try/0_Data_final/current/clip")
rlist=list.files(pattern="tif$",full.names = TRUE)
s_xvar=stack(rlist)
names(s_xvar)

#Projecting all to WGS_1984 because biomod doesn't work in NAD..
#r1=raster("D:/sdm/current_projection/Current_projection.grd")#reference for WGS projection
r1=raster("C:/Users/paude/OneDrive - Colostate/sdm/current_projection/Current_projection.grd")

r1

s_xvar1=projectRaster(s_xvar, crs = r1)#projecting stack downscaled variables along with soil
s_xvar1

#Testing multicollinearity 
AspenData <- read.csv("D:/BioMod2Try/Max_Aspen_Raster/samplingjune42023/presence.csv")#presence only data for multicollinearity
head(AspenData)
#Correlation Maxtix
library(dplyr)#for correlation matrix
presvals_all<-cbind(raster::extract(s_xvar1,AspenData[,c("longitude","lattitude")]),AspenData[,c("longitude","lattitude")])
cor_aspen_all<-presvals_all%>% select(-c("longitude","lattitude"))%>%cor()
cor_aspen_all
write.csv(cor_aspen_all,"D:/BioMod2Try/cor_newsamplewithpresenceonly_SRMClipped.csv")#saving correlation values
##for Variance Inflation Factor
library(usdm)
presvals_all<-cbind(raster::extract(s_xvar1,AspenData[,c("longitude","lattitude")]),AspenData[,c("longitude","lattitude")])
vifcor_1<-presvals_all%>% select(-c("longitude","lattitude"))%>%vifcor()
vifcor_1

#SDM Model_Response variable with 2000 presence and 2000 absence sample
##Response
AspenData <- read.csv("D:/BioMod2Try/Max_Aspen_Raster/samplingjune42023/AspenData.csv")
head(AspenData)
myRespName <- "Aspen"
myResp <- as.numeric(AspenData[,myRespName])
myRespXY <- AspenData[,c("longitude","lattitude")]

#Explanatory Variable and formatting for BioMod2
Explvar=raster::stack(s_xvar1$ADI,s_xvar1$MSPDD5,s_xvar1$MSP,s_xvar1$Clay,s_xvar1$pH,s_xvar1$om)
myExpl=raster::stack(Explvar)
myBiomodData <- BIOMOD_FormatingData(resp.var = myResp,
                                     expl.var = myExpl,
                                     resp.xy = myRespXY,
                                     resp.name = myRespName, )

myBiomodData
# Create default modeling options
myBiomodOption <- BIOMOD_ModelingOptions()
#Format response and predictor variable to biomod2 run setting
myBiomodData <- BIOMOD_FormatingData(resp.var = myResp,
                                     expl.var = myExpl,
                                     resp.xy = myRespXY,
                                     resp.name = myRespName, )
myBiomodData
                             
myBiomodOption <- BIOMOD_ModelingOptions()#this will include the default model evaluation metrics, ROC_AUC, Test Statistics which will help to understand model performance and pick algorithms that gives higher ROC-AUC value and greater TSS, The True Skill Statistic (TSS) is a measure of the accuracy of a binary classification model
Aspen_2010 <-
  BIOMOD_Modeling(
    bm.format = myBiomodData,
    bm.options = myBiomodOption,
    nb.rep = 10,
    data.split.perc = 70,
    var.import = 5,
    modeling.id = 'demo1'
  )
#Plot the evaluation metric to decide which algorithm to consider
bm_PlotEvalMean(Aspen_2010)#plots ROC-AUC and TSS
##Droped one model that had poor ROC-AUC and TSS
#Then Run Model Again dropping one algorith
Aspen_2010 <-
  BIOMOD_Modeling(
    bm.format = myBiomodData,
    models = c('GLM', 'GBM', 'GAM', 'CTA', 'ANN', 'MARS', 'RF', 'MAXENT.Phillips','FDA'),
    bm.options = myBiomodOption,
    nb.rep = 10,
    data.split.perc = 70,
    var.import = 5,
    modeling.id = 'demo1'
  )
#Now get importnace variables and evaluation metric as dataframe and save as CSV file
VarImport <- get_variables_importance(Aspen_2010)#gives mean of most important variables 
VarImport
apply(VarImport,c(1,2),mean)#mean of imp_variables
                
##Evaluation metrics in dataframe and save as CSV
myEval <- get_evaluations(Aspen_2010, as.data.frame = TRUE)
myEval
write.csv(myEval,"D:/BioMod2Try/0_Data_final/plots&results/results with newsample and clipped SRM/evalmetric.csv")#saving eval values
myEval$CV.strategy <- "Random"
myEval$CV.strategy[grepl("13", myEval$Model.name)] <- "Full"
myEval$CV.strategy[grepl("11|12", myEval$Model.name)] <- "Stratified"
head(myEval)
p=boxplot(myEval$Testing.data ~ interaction(myEval$Algo, myEval$CV.strategy),
        xlab = "", ylab = "ROC AUC", col = rep(c("brown", "cadetblue"), 3))
p=boxplot(myEval$Testing.data ~ interaction(myEval$Algo, myEval$CV.strategy),
          xlab = "", ylab = "ROC AUC")
p1=bm_PlotVarImpBoxplot(bm.out = Aspen_2010, group.by = c('expl.var', 'algo', 'algo'))

p2=bm_PlotVarImpBoxplot(bm.out = Aspen_2010, group.by = c('expl.var', 'algo', 'dataset'))
p3=bm_PlotVarImpBoxplot(bm.out = Aspen_2010, group.by = c('algo', 'expl.var', 'dataset'))

##Generate Response Curves
Aspen_rf<-BIOMOD_LoadModels(Aspen_2010,models="RF")

#Creating response curve
glm_eval_strip<-
  bm_PlotResponseCurves(
    bm.out=Aspen_2010,
    models=Aspen_rf,
    Data=get_formal_data(Aspen_2010,'expl.var'),
    show.variables=get_formal_data(Aspen_2010,'expl.var.names'),
    do.bivariate=FALSE,
    fixed.var.metric='median',
    legend=FALSE,
    display_title=FALSE,
    data_species=get_formal_data(Aspen_2010,'resp.var')
  )
###Now creating ensembled for better performance, the ensembled model is made selecting the algorithms that had better performance, like in our result we found the TSS hovers from 30%-57% therefore we considered 50% as the threshold and only those above 50% TSS is selected for ensembling. 
myBiomodEM <- BIOMOD_EnsembleModeling(bm.mod = Aspen_2010,
                                      models.chosen = 'all',
                                      em.by = 'all', 
                                      metric.select = c('TSS'),
                                      metric.select.thresh = c(0.5), 
                                      metric.eval = c('TSS', 'ROC'),
                                      var.import = 5,
                                      prob.mean = TRUE,
                                      prob.median = FALSE,
                                      prob.cv = FALSE,
                                      prob.ci = FALSE,
                                      prob.ci.alpha = 0.05,
                                      committee.averaging = TRUE,
                                      prob.mean.weight = TRUE,
                                      prob.mean.weight.decay = 'proportional',
                                      seed.val = 42)
#bm.mod = Aspen_2010,#specifies the model or dataset to be used in ensemble process
#models.chosen = 'all',#this indicates all available models will be used in Aspen_2010 
#em.by = 'all',#specifies that all available ensemble methods will be used 
#metric.select = c('TSS'),#specifies that the evaluation metric used for selecting the model is TSS, s a measure of the accuracy of a binary classification model
#metric.select.thresh = c(0.5),#sets threshold value for the selected evaluation metric Models with a TSS value greater than or equal to 0.5 will be considered for selection 
#metric.eval = c('TSS', 'ROC'),#Specifies the evaluation metric to be used for model evaluation. In addition to TSS, the Receiver Operating Characteristic (ROC) curve will also be evaluated
#var.import = 5, the number of top variables to be retained in the modeling process. Only top 5 with highest importance will be conidered. Usually 3-5 is recommended
#prob.mean = TRUE,#indicates that the mean probabilities will be calculated for the ensemble predictions
#prob.median = FALSE,#Median probabilities will not be calculated
#prob.cv = FALSE,# cross validation will not be used to estimate probabilities
#prob.ci = FALSE,#confidence interval for the probabilities will not be calculated
#prob.ci.alpha = 0.05,if prob.ci was set to true then this defines alpha level (significance level) for the confidence interval
#committee.averaging = TRUE,#indicates committee averaging will be used to aggregate the predictions from different models in the ensemble
#prob.mean.weight = TRUE,# mean weight will be calculated for the ensemble models
#prob.mean.weight.decay = 'proportional',proportional decay method for mean weights means model performing better are asigned higer weight and vice-versa.
#seed.val = 42)#this parameter sets the value for reproducibility, The same seed value will produce the same results if the process is run again
p4=bm_PlotVarImpBoxplot(bm.out = myBiomodEM, group.by = c('expl.var', 'algo', 'algo'))
p5=bm_PlotVarImpBoxplot(bm.out = myBiomodEM, group.by = c('expl.var', 'algo', 'dataset'))
p6=bm_PlotVarImpBoxplot(bm.out = myBiomodEM, group.by = c('algo', 'expl.var', 'dataset'))

#Get Evaluation Metric of Ensembled Models
myEval <- get_evaluations(myBiomodEM, as.data.frame = TRUE)
myEval


####Ensembled varimportnace
VarImport <- get_variables_importance(myBiomodEM)#gives mean of most important variables 
VarImport
apply(VarImport,c(1,2),mean)#mean of imp_variables
p4=bm_PlotVarImpBoxplot(bm.out = myBiomodEM, group.by = c('expl.var', 'algo', 'algo'))
p5=bm_PlotVarImpBoxplot(bm.out = myBiomodEM, group.by = c('expl.var', 'algo', 'dataset'))
p6=bm_PlotVarImpBoxplot(bm.out = myBiomodEM, group.by = c('algo', 'expl.var', 'dataset'))
#Projecting from single models
myBiomodProj <- BIOMOD_Projection(bm.mod = Aspen_2010,#invidual models
                                  proj.name = 'Current',
                                  new.env = myExpl,
                                  models.chosen = 'all',
                                  metric.binary = 'all',
                                  metric.filter = 'all',
                                  build.clamping.mask = TRUE,
                                  output.format = '.img')
myBiomodProj

plot(myBiomodProj)
##Projecting from ensembled models 
BiomodEMProj <- BIOMOD_EnsembleForecasting(bm.em = myBiomodEM,#consider ensembled model
                                           proj.name = 'CurrentEM',
                                           new.env = myExpl,
                                           models.chosen = 'all',
                                           metric.binary = 'all',
                                           metric.filter = 'all',
                                           output.format = '.img')

BiomodEMProj
plot(BiomodEMProj)

CurrentProj <- get_predictions(myBiomodProj)
CurrentProj


##Climate Change Scenario SS2_4.5_2010_2040
setwd("D:/BioMod2Try/0_Data_final/LowEmisSce/2040/clip")
#setwd("C:/Users/paude/OneDrive - Colostate/sdm/0_downscaled_variables/Ss4.5_2041_2070")
rlist=list.files(pattern="tif$",full.names = TRUE)
s_xvar=stack(rlist)
names(s_xvar)
#r1=raster("D:/sdm/current_projection/Current_projection.grd")#reference for WGS projection
r1=raster("C:/Users/paude/OneDrive - Colostate/sdm/current_projection/Current_projection.grd")

r1

s_xvar1=projectRaster(s_xvar, crs = r1)#projecting stack downscaled variables along with soil
s_xvar1

#2040
Explvar_2040=raster::stack(s_xvar1$ADI,s_xvar1$MSPDD5,s_xvar1$Clay,s_xvar1$pH,s_xvar1$MSP,s_xvar1$om )
#Explvar=raster::stack(s_xvar1$ADI1,s_xvar1$GSPDD5,s_xvar1$RH,s_xvar1$MSP,s_xvar1$hli,s_xvar1$clay,s_xvar1$ph)

#Explvar=raster::stack(s_xvar1$ADI1,s_xvar1$GSPDD5,s_xvar1$RH,s_xvar1$TD)

myExpl_2040=raster::stack(Explvar_2040)


myBiomodData_2040 <- BIOMOD_FormatingData(resp.var = myResp,
                                     expl.var = myExpl_2040,
                                     resp.xy = myRespXY,
                                     resp.name = myRespName, )
myBiomodData_2040
#plot(myBiomodData_2040)                               
myBiomodOption <- BIOMOD_ModelingOptions()
Aspen_2040 <-
  BIOMOD_Modeling(
    bm.format = myBiomodData_2040,
    models = c('GLM', 'GBM', 'GAM', 'CTA', 'ANN', 'MARS', 'RF', 'MAXENT.Phillips','FDA'),
    bm.options = myBiomodOption,
    nb.rep = 10,
    data.split.perc = 70,
    var.import = 5,
    modeling.id = 'demo1'
  )
Aspen_2040

myBiomodEM_2040 <- BIOMOD_EnsembleModeling(bm.mod = Aspen_2040,
                                      models.chosen = 'all',
                                      em.by = 'all',
                                      metric.select = c('TSS'),
                                      metric.select.thresh = c(0.5),
                                      metric.eval = c('TSS', 'ROC'),
                                      var.import = 5,
                                      prob.mean = TRUE,
                                      prob.median = FALSE,
                                      prob.cv = FALSE,
                                      prob.ci = FALSE,
                                      prob.ci.alpha = 0.05,
                                      committee.averaging = TRUE,
                                      prob.mean.weight = TRUE,
                                      prob.mean.weight.decay = 'proportional',
                                      seed.val = 42)
myBiomodEM_2040

BiomodEMProj_2040 <- BIOMOD_EnsembleForecasting(bm.em = myBiomodEM_2040,
                                           proj.name = 'BiomodEMProj_2040',
                                           new.env = myExpl_2040,
                                           models.chosen = 'all',
                                           metric.binary = 'all',
                                           metric.filter = 'all',
                                           output.format = '.img')


FutureProj_2040 <- get_predictions(BiomodEMProj_2040)

FutureProj_2040

##Climate Change Scenario SS2_4.5_2041_2070
setwd("D:/BioMod2Try/0_Data_final/LowEmisSce/2070/clip")
#setwd("C:/Users/paude/OneDrive - Colostate/sdm/0_downscaled_variables/Ss4.5_2041_2070")
rlist=list.files(pattern="tif$",full.names = TRUE)
s_xvar=stack(rlist)
names(s_xvar)
#r1=raster("D:/sdm/current_projection/Current_projection.grd")#reference for WGS projection
r1=raster("C:/Users/paude/OneDrive - Colostate/sdm/current_projection/Current_projection.grd")

r1

s_xvar1=projectRaster(s_xvar, crs = r1)#projecting stack downscaled variables along with soil
s_xvar1  
####2070
Explvar_2070=raster::stack(s_xvar1$ADI,s_xvar1$MSPDD5,s_xvar1$MSP,s_xvar1$Clay,s_xvar1$pH,s_xvar1$om)
#Explvar=raster::stack(s_xvar1$ADI1,s_xvar1$GSPDD5,s_xvar1$RH,s_xvar1$MSP,s_xvar1$hli,s_xvar1$clay,s_xvar1$ph)

#Explvar=raster::stack(s_xvar1$ADI1,s_xvar1$GSPDD5,s_xvar1$RH,s_xvar1$TD)

myExpl_2070=raster::stack(Explvar_2070)


myBiomodData_2070 <- BIOMOD_FormatingData(resp.var = myResp,
                                          expl.var = myExpl_2070,
                                          resp.xy = myRespXY,
                                          resp.name = myRespName, )
myBiomodData_2070
plot(myBiomodData_2070)                               
myBiomodOption <- BIOMOD_ModelingOptions()
Aspen_2070 <-
  BIOMOD_Modeling(
    bm.format = myBiomodData_2070,
    models = c('GLM', 'GBM', 'GAM', 'CTA', 'ANN', 'MARS', 'RF', 'MAXENT.Phillips','FDA'),
    bm.options = myBiomodOption,
    nb.rep = 10,
    data.split.perc = 70,
    var.import = 5,
    modeling.id = 'demo1'
  )
Aspen_2070

myBiomodEM_2070 <- BIOMOD_EnsembleModeling(bm.mod = Aspen_2070,
                                           models.chosen = 'all',
                                           em.by = 'all',
                                           metric.select = c('TSS'),
                                           metric.select.thresh = c(0.5),
                                           metric.eval = c('TSS', 'ROC'),
                                           var.import = 5,
                                           prob.mean = TRUE,
                                           prob.median = FALSE,
                                           prob.cv = FALSE,
                                           prob.ci = FALSE,
                                           prob.ci.alpha = 0.05,
                                           committee.averaging = TRUE,
                                           prob.mean.weight = TRUE,
                                           prob.mean.weight.decay = 'proportional',
                                           seed.val = 42)


myBiomodEM_2070
BiomodEMProj_2070 <- BIOMOD_EnsembleForecasting(bm.em = myBiomodEM_2070,
                                                proj.name = 'BiomodEMProj_2070',
                                                new.env = myExpl_2070,
                                                models.chosen = 'all',
                                                metric.binary = 'all',
                                                metric.filter = 'all',
                                                output.format = '.img')

FutureProj_2070 <- get_predictions(BiomodEMProj_2070)
FutureProj_2070

##2100
##Climate Change Scenario SS2_4.5_2071_2100
#C:/Users/paude/Desktop/BioMod2Try/SS245_biomod2/Ssp4.5_2071_2100
setwd("D:/BioMod2Try/0_Data_final/LowEmisSce/2100/clip")
#setwd("C:/Users/paude/OneDrive - Colostate/sdm/0_downscaled_variables/Ss4.5_2041_2070")
rlist=list.files(pattern="tif$",full.names = TRUE)
s_xvar=stack(rlist)
names(s_xvar)
#r1=raster("D:/sdm/current_projection/Current_projection.grd")#reference for WGS projection
r1=raster("C:/Users/paude/OneDrive - Colostate/sdm/current_projection/Current_projection.grd")

r1

s_xvar1=projectRaster(s_xvar, crs = r1)#projecting stack downscaled variables along with soil
s_xvar1  
####2100
Explvar_2100=raster::stack(s_xvar1$ADI,s_xvar1$MSPDD5,s_xvar1$MSP,s_xvar1$Clay,s_xvar1$pH, s_xvar1$om)
#Explvar=raster::stack(s_xvar1$ADI1,s_xvar1$GSPDD5,s_xvar1$RH,s_xvar1$MSP,s_xvar1$hli,s_xvar1$clay,s_xvar1$ph)

#Explvar=raster::stack(s_xvar1$ADI1,s_xvar1$GSPDD5,s_xvar1$RH,s_xvar1$TD)

myExpl_2100=raster::stack(Explvar_2100)


myBiomodData_2100 <- BIOMOD_FormatingData(resp.var = myResp,
                                          expl.var = myExpl_2100,
                                          resp.xy = myRespXY,
                                          resp.name = myRespName, )
myBiomodData_2100
plot(myBiomodData_2100)                               
myBiomodOption_2100 <- BIOMOD_ModelingOptions()
Aspen_2100 <-
  BIOMOD_Modeling(
    bm.format = myBiomodData_2100,
    models = c('GLM', 'GBM', 'GAM', 'CTA', 'ANN', 'MARS', 'RF', 'MAXENT.Phillips','FDA'),
    bm.options = myBiomodOption,
    nb.rep = 10,
    data.split.perc = 70,
    var.import = 5,
    modeling.id = 'demo1'
  )
Aspen_2100

myBiomodEM_2100 <- BIOMOD_EnsembleModeling(bm.mod = Aspen_2100,
                                           models.chosen = 'all',
                                           em.by = 'all',
                                           metric.select = c('TSS'),
                                           metric.select.thresh = c(0.5),
                                           metric.eval = c('TSS', 'ROC'),
                                           var.import = 5,
                                           prob.mean = TRUE,
                                           prob.median = FALSE,
                                           prob.cv = FALSE,
                                           prob.ci = FALSE,
                                           prob.ci.alpha = 0.05,
                                           committee.averaging = TRUE,
                                           prob.mean.weight = TRUE,
                                           prob.mean.weight.decay = 'proportional',
                                           seed.val = 42)



BiomodEMProj_2100 <- BIOMOD_EnsembleForecasting(bm.em = myBiomodEM_2100,
                                                proj.name = 'BiomodEMProj_2100',
                                                new.env = myExpl_2100,
                                                models.chosen = 'all',
                                                metric.binary = 'all',
                                                metric.filter = 'all',
                                                output.format = '.img')

BiomodEMProj_2100
plot(BiomodEMProj_2100)

FutureProj_2100 <- get_predictions(BiomodEMProj_2100)
FutureProj_2100


CurrentProj <- get_predictions(BiomodEMProj)
CurrentProj

###Species Range Change 2011-2040####

SRC_Curr2040 <- BIOMOD_RangeSize(proj.current = CurrentProj, proj.future = FutureProj_2040)
SRC_Curr2040

SRC_Curr2040$Compt.By.Models

###Species Range Change 2041-2070

SRC_Curr2070 <- BIOMOD_RangeSize(proj.current = FutureProj_2040, proj.future = FutureProj_2070)
SRC_Curr2070

SRC_Curr2070$Compt.By.Models

###Species Range Change 2071-2100

SRC_Curr2100 <- BIOMOD_RangeSize(proj.current = FutureProj_2070, proj.future = FutureProj_2100)
SRC_Curr2100

SRC_Curr2100$Compt.By.Models

######Species Range Change 2011-2100

SRC_Curr2011_2100 <- BIOMOD_RangeSize(proj.current = CurrentProj, proj.future = FutureProj_2100)
SRC_Curr2011_2100

SRC_Curr2011_2100$Compt.By.Models


######Species Range Change 2011-2070

SRC_Curr2011_2070 <- BIOMOD_RangeSize(proj.current = CurrentProj, proj.future = FutureProj_2070)
SRC_Curr2011_2070

SRC_Curr2011_2070$Compt.By.Models

######Species Range Change 2011-2040

SRC_Curr2011_2040 <- BIOMOD_RangeSize(proj.current = CurrentProj, proj.future = FutureProj_2040)
SRC_Curr2011_2040

SRC_Curr2011_2040$Compt.By.Models


###Load binary projection
Aspen_bin_proj_current<-
  stack(
    c(
      ca="D:/BioMod2Try/0_Data_final/current/clip/Aspen/proj_CurrentEM/individual_projections/Aspen_EMcaByTSS_mergedAlgo_mergedRun_mergedData.img",
      wm="D:/BioMod2Try/0_Data_final/current/clip/Aspen/proj_CurrentEM/individual_projections/Aspen_EMwmeanByTSS_mergedAlgo_mergedRun_mergedData.img"
    )
  )


##2040
Aspen_bin_proj_2040<-
  stack(
    c(
      ca="D:/BioMod2Try/0_Data_final/LowEmisSce/2040/clip/Aspen/proj_BiomodEMProj_2040/individual_projections/Aspen_EMcaByTSS_mergedAlgo_mergedRun_mergedData.img",
      wm="D:/BioMod2Try/0_Data_final/LowEmisSce/2040/clip/Aspen/proj_BiomodEMProj_2040/individual_projections/Aspen_EMwmeanByTSS_mergedAlgo_mergedRun_mergedData.img"
    )
  )

##2070
Aspen_bin_proj_2070<-
  stack(
    c(
      ca="D:/BioMod2Try/0_Data_final/LowEmisSce/2070/clip/Aspen/proj_BiomodEMProj_2070/individual_projections/Aspen_EMcaByTSS_mergedAlgo_mergedRun_mergedData.img",
      wm="D:/BioMod2Try/0_Data_final/LowEmisSce/2070/clip/Aspen/proj_BiomodEMProj_2070/individual_projections/Aspen_EMwmeanByTSS_mergedAlgo_mergedRun_mergedData.img"
    )
  )

###2100
Aspen_bin_proj_2100<-
  stack(
    c(
      ca="D:/BioMod2Try/0_Data_final/LowEmisSce/2100/clip/Aspen/proj_BiomodEMProj_2100/individual_projections/Aspen_EMcaByTSS_mergedAlgo_mergedRun_mergedData.img",
      wm="D:/BioMod2Try/0_Data_final/LowEmisSce/2100/clip/Aspen/proj_BiomodEMProj_2100/individual_projections/Aspen_EMwmeanByTSS_mergedAlgo_mergedRun_mergedData.img"
    )
  )

############
SRC_2011_2040<-BIOMOD_RangeSize(
  Aspen_bin_proj_current,
  Aspen_bin_proj_2040
)

SRC_2011_2040$Compt.By.Models


###2011-2070
SRC_2011_2070<-BIOMOD_RangeSize(
  Aspen_bin_proj_current,
  Aspen_bin_proj_2070
)

SRC_2011_2070$Compt.By.Models

###2011-2100
SRC_2011_2100<-BIOMOD_RangeSize(
  Aspen_bin_proj_current,
  Aspen_bin_proj_2100
)

SRC_2011_2100$Compt.By.Models

##2040-2070
SRC_2040_2070<-BIOMOD_RangeSize(
  Aspen_bin_proj_2040,
  Aspen_bin_proj_2070
)

SRC_2040_2070$Compt.By.Models
###
AspenSrcMap<-
  stack(
    SRC_2011_2040$Diff.By.Pixel,
    SRC_2011_2070$Diff.By.Pixel,
    SRC_2011_2100$Diff.By.Pixel
  )
names(AspenSrcMap)<-c("ca cur_2040",
                      "wm cur_2040",
                      "ca cur-2070",
                      "wm cur-2070",
                      "ca cur-2100",
                      "wm cur-2100"
                      )

my.at<-seq(-2.5,1.5,1)
mycolorkey<-
  list(
    at=my.at,#where the color change
    labels=
      list(
        labels=c("lost","pres","abs","gain"),
        at=my.at[-1]-0.5
      )
  )
rasterVis::levelplot(
  AspenSrcMap,
  main="Aspen Range Change",
  colorkey=mycolorkey,
  col.regions=c("#f03b20","#99d8c9","#f0f0f0","#2ca25f"),
  layout=c(2,3)
)


```

